---
title: "Angle Array Calculations"
output: html_notebook
---

**Purpose**

The purpose of this notebook is to calculate the angle arrays for each joint using the quaternions from the cleaned mocap data. 

```{r}

# import the necessary libraries
library(readr)
library(pracma)
library(dplyr)
library(tidyr)
library(data.table)

```

```{r}

raise_rhand_mocap = read.csv("/Users/teresanguyen/Documents/Faboratory Stuff/soft-robotics-simulations/data/raise_rhand_mocap_cleaned.csv")
raise_rhand_sensor = read.csv("/Users/teresanguyen/Documents/Faboratory Stuff/soft-robotics-simulations/data/raise_rhand_sensor.csv")

colnames(raise_rhand_mocap)
head(raise_rhand_mocap)

```


```{r}

sensor_dt <- as.data.table(raise_rhand_sensor)

setnames(sensor_dt, "Time..ms.", "time_ms")

start_time <- sensor_dt$time_ms[1]
sensor_dt[, time_ms := time_ms - start_time]

sensor_dt[, time_sec := time_ms / 1000]

head(sensor_dt[, .(time_ms, time_sec)])
head(sensor_dt)

```

```{r}

headers <- c("time_ms", "x", "y", "z", "qw", "qx", "qy", "qz",
             "abdomen_z", "abdomen_y", "abdomen_x",
             "hip_x_right", "hip_z_right", "hip_y_right", "knee_right", "ankle_y_right", "ankle_x_right",
             "hip_x_left", "hip_z_left", "hip_y_left", "knee_left", "ankle_y_left", "ankle_x_left",
             "shoulder1_right", "shoulder2_right", "elbow_right",
             "shoulder1_left", "shoulder2_left", "elbow_left")

# units <- c("",
#            "rad", "rad", "rad",
#            "rad", "rad", "rad", "rad", "rad", "rad",
#            "rad", "rad", "rad", "rad", "rad", "rad",
#            "rad", "rad", "rad", "rad", "rad", "rad")

# headers_with_units <- ifelse(units == "", headers, paste0(headers, " [", units, "]"))

n_rows <- nrow(raise_rhand_mocap)
angle_array <- data.frame(matrix(NA, nrow = n_rows, ncol = length(headers)))
colnames(angle_array) <- headers

angle_array$time_ms <- raise_rhand_mocap$time_ms

head(angle_array)

```

```{r}

quat_to_euler <- function(qx, qy, qz, qw) {

  # roll (x)
  sinr_cosp <- 2 * (qw * qx + qy * qz)
  cosr_cosp <- 1 - 2 * (qx^2 + qy^2)
  roll <- atan2(sinr_cosp, cosr_cosp)

  # pitch (y)
  sinp <- 2 * (qw * qy - qz * qx)
  pitch <- ifelse(abs(sinp) >= 1, sign(sinp) * pi/2, asin(sinp))

  # yaw (z)
  siny_cosp <- 2 * (qw * qz + qx * qy)
  cosy_cosp <- 1 - 2 * (qy^2 + qz^2)
  yaw <- atan2(siny_cosp, cosy_cosp)

  return(c(roll, pitch, yaw))
}

```

**Base**

```{r}

angle_array$x <- raise_rhand_mocap$`base_X` / 1000
angle_array$y <- raise_rhand_mocap$`base_Y.1` / 1000
angle_array$z <- raise_rhand_mocap$`base_Z.1` / 1000
angle_array$qw <- raise_rhand_mocap$`base_W`
angle_array$qx <- raise_rhand_mocap$`Skeleton.002.Skeleton.002_X`
angle_array$qy <- raise_rhand_mocap$`base_Y`
angle_array$qz <- raise_rhand_mocap$`base_Z`

head(angle_array)

```

**Abdomen**

```{r}

quats <- raise_rhand_mocap %>% select(time_ms, qx = abdomen_X, qy = abdomen_Y, qz = abdomen_Z, qw = abdomen_W)

euler_matrix <- t(mapply(quat_to_euler, quats$qx, quats$qy, quats$qz, quats$qw))
euler_df <- as.data.frame(euler_matrix)
colnames(euler_df) <- c("abdomen_x", "abdomen_y", "abdomen_z")

angle_array$`abdomen_z` <- euler_df$`abdomen_z`
angle_array$`abdomen_y` <- euler_df$`abdomen_y`
angle_array$`abdomen_x` <- euler_df$`abdomen_x`

head(angle_array)

```

**Hip Right**

```{r}

quats_hip <- raise_rhand_mocap %>% select(qx = right_thigh_X, qy = right_thigh_Y, qz = right_thigh_Z, qw = right_thigh_W)

euler_matrix_hip <- t(mapply(quat_to_euler, quats_hip$qx, quats_hip$qy, quats_hip$qz, quats_hip$qw))
euler_df_hip <- as.data.frame(euler_matrix_hip)
colnames(euler_df_hip) <- c("hip_x_right", "hip_y_right", "hip_z_right")

angle_array$`hip_x_right` <- euler_df_hip$`hip_x_right`
angle_array$`hip_z_right` <- euler_df_hip$`hip_z_right`
angle_array$`hip_y_right` <- euler_df_hip$`hip_y_right`

head(angle_array[, c("time_ms", "hip_x_right", "hip_y_right", "hip_z_right")])


head(angle_array)
# "left_thigh_X" %in% colnames(raise_rhand_mocap)


```

**Knee Right**

```{r}

quats_knee_right <- raise_rhand_mocap %>% select(qx = right_shin_X, qy = right_shin_Y, qz = right_shin_Z, qw = right_shin_W)

euler_knee_right <- t(mapply(quat_to_euler, quats_knee_right$qx, quats_knee_right$qy, quats_knee_right$qz, quats_knee_right$qw))
knee_right_angle <- as.data.frame(euler_knee_right)[,1]

angle_array$`knee_right` <- knee_right_angle

head(angle_array)

```

**Ankle Right**

```{r}

quats_ankle_right <- raise_rhand_mocap %>% select(qx = right_foot_X, qy = right_foot_Y, qz = right_foot_Z, qw = right_foot_W)

euler_ankle_right <- t(mapply(quat_to_euler, quats_ankle_right$qx, quats_ankle_right$qy, quats_ankle_right$qz, quats_ankle_right$qw))
euler_df_ankle_right <- as.data.frame(euler_ankle_right)

angle_array$`ankle_x_right` <- euler_df_ankle_right[, 1]
angle_array$`ankle_y_right` <- euler_df_ankle_right[, 2]

head(angle_array[, c("time_ms", "ankle_x_right", "ankle_y_right")])

head(angle_array)


```

**Hip Left**

```{r}

quats_hip <- raise_rhand_mocap %>% select(qx = left_thigh_X, qy = left_thigh_Y, qz = left_thigh_Z, qw = left_thigh_W)
colnames(raise_rhand_mocap)
euler_matrix_hip <- t(mapply(quat_to_euler, quats_hip$qx, quats_hip$qy, quats_hip$qz, quats_hip$qw))
euler_df_hip <- as.data.frame(euler_matrix_hip)
colnames(euler_df_hip) <- c("hip_x_left [rad]", "hip_y_left [rad]", "hip_z_left [rad]")

angle_array$`hip_x_left` <- euler_df_hip$`hip_x_left [rad]`
angle_array$`hip_z_left` <- euler_df_hip$`hip_z_left [rad]`
angle_array$`hip_y_left` <- euler_df_hip$`hip_y_left [rad]`

head(angle_array[, c("time_ms", "hip_x_left", "hip_z_left", "hip_y_left")])

head(angle_array)
colnames(angle_array)

```

**Knees**

For knees, we only keep 1 angle (roll - X) because the main rotation happens around the X axis. 
"No, you don't need to control all three angles (flexion, extension, and rotation) in the knee joint for typical daily activities. The knee primarily functions with flexion and extension (bending and straightening). While there's some degree of rotational movement, it's minimal and primarily occurs with the ankle and hip joints."

**Knee Left**

```{r}

quats_knee_left <- raise_rhand_mocap %>% select(qx = left_shin_X, qy = left_shin_Y, qz = left_shin_Z, qw = left_shin_W)

euler_knee_left <- t(mapply(quat_to_euler, quats_knee_left$qx, quats_knee_left$qy, quats_knee_left$qz, quats_knee_left$qw))
knee_left_angle <- as.data.frame(euler_knee_left)[,1]

angle_array$`knee_left` <- knee_left_angle

head(angle_array)
colnames(angle_array)

```

Ankles are primarily focused on the plantar and dorsiflexion (X) and inversion eversion (Y) 


**Ankle Left**

```{r}

quats_ankle_left <- raise_rhand_mocap %>% select(qx = left_foot_X, qy = left_foot_Y, qz = left_foot_Z, qw = left_foot_W)

euler_ankle_left <- t(mapply(quat_to_euler, quats_ankle_left$qx, quats_ankle_left$qy, quats_ankle_left$qz, quats_ankle_left$qw))
euler_df_ankle_left <- as.data.frame(euler_ankle_left)

angle_array$`ankle_y_left` <- euler_df_ankle_left[, 2]
angle_array$`ankle_x_left` <- euler_df_ankle_left[, 1]

head(angle_array)

```

**Shoulders**

We use upper arms instead of shoulders from the mocap data because the shoulder joint controls the movement of the uper arm so in order to get the shoulder joint movements, we need to analuze how the upper arm is oriented over time.

***Right Shoulder***

```{r}

quats_shoulder_right <- raise_rhand_mocap %>% select(qx = right_upper_arm_X, qy = right_upper_arm_Y, qz = right_upper_arm_Z, qw = right_upper_arm_W)
euler_shoulder_right <- t(mapply(quat_to_euler, quats_shoulder_right$qx, quats_shoulder_right$qy, quats_shoulder_right$qz, quats_shoulder_right$qw))

euler_df_shoulder_right <- as.data.frame(euler_shoulder_right)

angle_array$`shoulder1_right` <- euler_df_shoulder_right[, 2]
angle_array$`shoulder2_right` <- euler_df_shoulder_right[, 1]

head(angle_array)

```

**Elbow Right**

```{r}
quats_elbow_right <- raise_rhand_mocap %>% select(qx = right_forearm_X, qy = right_forearm_Y, qz = right_forearm_Z, qw = right_forearm_W)
euler_df_elbow_right <- as.data.frame(t(mapply(quat_to_euler, quats_elbow_right$qx, quats_elbow_right$qy, quats_elbow_right$qz, quats_elbow_right$qw)))
angle_array$`elbow_right` <- euler_df_elbow_right[, 1]

head(angle_array)

```

**Shoulder Left**

```{r}

quats_shoulder_left <- raise_rhand_mocap %>% select(qx = left_upper_arm_X, qy = left_upper_arm_Y, qz = left_upper_arm_Z, qw = left_upper_arm_W)

euler_shoulder_left <- t(mapply(quat_to_euler, quats_shoulder_left$qx, quats_shoulder_left$qy, quats_shoulder_left$qz,
                                quats_shoulder_left$qw))
euler_df_shoulder_left <- as.data.frame(euler_shoulder_left)

angle_array$`shoulder1_left` <- euler_df_shoulder_left[, 2]
angle_array$`shoulder2_left` <- euler_df_shoulder_left[, 1]

head(angle_array[, c("time_ms", "shoulder1_left", "shoulder2_left")])

```

***Elbow Left***

```{r}

quats_elbow_left <- raise_rhand_mocap %>% select(qx = left_forearm_X, qy = left_forearm_Y, qz = left_forearm_Z, qw = left_forearm_W)
euler_df_elbow_left <- as.data.frame(t(mapply(quat_to_euler, quats_elbow_left$qx, quats_elbow_left$qy, quats_elbow_left$qz, quats_elbow_left$qw)))
angle_array$`elbow_left` <- euler_df_elbow_left[, 1]

head(angle_array)

```


```{r}

path <- file.path("~", "Documents", "Faboratory Stuff", "soft-robotics-simulations", "data", "angle_array_rhand_mocap.csv")
write.csv(angle_array, path, row.names = FALSE)

```
